---
title: "Tutorial on cheese"
author: Alex Zajichek
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output: 
    rmarkdown::html_vignette
vignette: >
    %\VignetteIndexEntry{Tutorial on cheese}
    %\VignetteEngine{knitr::rmarkdown}
    \usepackage[utf8]{inputenc}
---

#Introduction
The data used in the following examples comes from the heart disease dataset found at the [UCI Machine Learning Repository](http://archive.ics.uci.edu/ml/datasets/Heart+Disease).

```{r}
#Load packages
require(tidyverse); require(cheese)

#Look at the top ten rows
heart_disease

```

#Creating a univariate table
The function `univariate_table` allows flexible summarization and presentation of variables in a dataset. Arguments are available to customize the statistics that are computed, association metrics, stratification variables, variable labels, etc. The `format` argument allows the user to render any table in "html", "latex", "markdown", "pandoc", "none" (i.e. return a `data.frame`). The following examples are rendered in "html" (default):

##Default
By default, the median (iqr), count (%), and the number of distinct values are displayed for numeric, categorical, and 'other' data types, 
```{r}
#Default table
heart_disease %>%
    univariate_table
```



